---
title: "Practice: Data Wrangling"
format: pdf
---

## Instructions: starting code

In a Quarto document file, run the following commands:

```{r message=FALSE}
# packages
library(tidyverse)
```


```{r}
students = data.frame(
  "name" = c("Harry", "Ron", "Hermione", "Luna", "Draco"),
  "gender" = c("male", "male", "female", "female", "male"),
  "gpa" = c(3.8, 3.4, 4, 3.5, 3.9),
  "major" = c("econ", "cogsci", "stats", "psych", "mcb"),
  "transfer" = c(TRUE, FALSE, FALSE, TRUE, FALSE)
)

students
```


## Data wrangling functions (reminder):

- `slice()`: subset data based on rows position
- `filter()`: subset data based on a logical condition
- `select()`: select columns by name or index
- `arrange()`: arrange content on one or more variables
- `count()`: calculate counts
- `mutate()`: change an existing column, or add a new column
- `summarize()`: calculate numeric summaries
- `group_by()`: group data based on one or more columns


## 1) Your Turn: Fill in the blanks

Each of the following problems has a code snippet with an incomplete command. 
Use the code chunks to write your answers, i.e. use the provided command and
fill in the blanks.

\vspace{5mm}

```r
# 1) use slice to subset the first row
slice(_______, ___)
```

```{r}
# 1) Answer

```

\vspace{5mm}

```r
# 2) use slice to subset the first three rows
# hint: the colon operator ':' is your friend
slice(_______, ____)
```

```{r}
# 2) Answer

```

\vspace{5mm}

```r
# 3) slice students to get the data of the 1st, 3rd, and 5th rows
# hint: the combine function c() is your friend
slice(________, _______)
```

```{r}
# 3) Answer

```

\vspace{5mm}

```r
# 4) select columns name and major
select(________, ________, _______)
```

```{r}
# 4) Answer

```

\vspace{5mm}

```r
# 5) select columns in 2nd, 3rd and 4th positions
# hint: the colon operator ':' can be your friend
select(________, _____)
```

```{r}
# 5) Answer

```

\vspace{5mm}

```r
# 6) create a data frame "women" by filtering female students
# hint: the == operator is your friend
women <- filter(________, __________________)
women
```

```{r}
# 6) Answer

```

\vspace{5mm}

```r
# 7) create a data frame "men" by filtering students that are not female
# hint: the != operator is your friend
men <- filter(________, __________________)
men
```

```{r}
# 7) Answer

```

\vspace{5mm}

```r
# 8) subset students with gpa greater than 3.5
filter(_______, __________)
```

```{r}
# 8) Answer

```

\vspace{5mm}

```r
# 9) subset male students with gpa greater than or equal to 3.6
# hint: the & operator is your friend
filter(_______, ______________________)
```

```{r}
# 9) Answer

```

\vspace{5mm}

```r
# 10) subset students majoring in econ and stats
# hint: the %in% operator is your friend
filter(________, __________________________)
```

```{r}
# 10) Answer

```

\vspace{5mm}

```r
# 11) subset female students OR students majoring in cogsci
# hint: the | operator is your friend
filter(________, __________________________)
```

```{r}
# 11) Answer

```

\vspace{5mm}

```r
# 12) arrange students by gpa values (in ascending order)
arrange(________, ____)
```

```{r}
# 12)Answer

```

\vspace{5mm}

```r
# 13) arrange students by gpa values (in descending order)
arrange(________, _________)
```

```{r}
# 13) Answer

```

\vspace{5mm}

```r
# 14) arrange students by their names (in alphabetical order)
arrange(________, _____)
```

```{r}
# 14) Answer

```

\vspace{5mm}

```r
# 15) arrange students by gender, and gpa
arrange(________, _______, _____)
```

```{r}
# 15) Answer

```

\vspace{5mm}

```r
# 16) use mutate() to create a data frame "students2" 
# by adding a column gpa_sqr, squaring the gpa values
students2 = mutate(_______, gpa_sqr = _________)
students2
```

```{r}
# 16) Answer

```

\vspace{5mm}


## 2) Your Turn: No More Training Wheels

Let's see if you can answer the following problems without us providing an
incomplete command.

17 - Use summarize to compute the mean gpa (of all students)

```{r}
# 17) Answer

```

\vspace{5mm}

18 - Compute mean gpa of female students

```{r}
# 18) Answer

```

\vspace{5mm}

19 - Compute min and max gpa of male students

```{r}
# 19) Answer

```

\vspace{5mm}

20 - Compute median gpa, grouped by gender

```{r}
# 20) Answer

```

\vspace{5mm}

21 - Compute mean and median gpa, grouped by transfer status

```{r}
# 21) Answer

```

\vspace{5mm}

22 - Obtain a table of male students, that are not transfer, arranging its 
contents by name in alphabetical order (A-Z), selecting columns name, 
major, and gpa (in that order) 

```{r}
# 22) Answer

```
